Ecosystem functional response across precipitation extremes in a sagebrush steppe

### 2018-01-05 manuscript/usses_water_ms.Rmd
big update with plug in VWC values from SOILWAT

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artemesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
During the period of our experiment (2011 -- 2015), average mean annual precipitation was 265 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

@@

Ambient precipitation was variable over the five years of the study (Fig. 2A), meaning our experiment captured a range of environmental conditions.
ANPP varied from a minimum of 74.5 g m$^{-2}$ in 2014 to a maximum of 237.1 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
ANPP was slightly higher in irrigation plots \new{(on average 17\% higher)} and slightly lower in drought plots \new{(on average 23\% lower) relative to control plots} (Fig. 2C), corresponding to estimated soil volumetric water content (VWC) differences among treatments (Fig. 2B).
\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
Such differences in soil VWC indicate our treatment infrastructure was successful.
ANPP was highly variable across plots within years, \new{as indicated by the large and overlapping standard deviations} (Fig. 2C).

@@

\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0002; Fig. 3C)}.

>>>>

We conducted our precipitation manipulation experiment in a sagebrush steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near Dubois, Idaho (44.2$^{\circ}$ N, 112.1$^{\circ}$ W), 1500 m above sea level.
The plant community is dominated by the shrub \emph{Artemesia tripartita}, \new{the perennial forb \emph{Balsamorhiza sagittata}}, and three perennial bunchgrasses, \emph{Pseudoroegneria spicata}, \emph{Poa secunda}, and \emph{Hesperostipa comata} \new{(see Appendix 1 for rank abundance curves)}.
During the period of our experiment (2011 -- 2016), mean annual precipitation was 250 mm year$\phantom{}^{-1}$ and mean monthly temperature ranged from -5.2$^{\circ}$C in January to 21.8$^{\circ}$C in July.
Between 1926 and 1932, range scientists at the USSES established 26 permanent 1 m$^2$ quadrats to track vegetation change over time.
In 2007, we relocated 14 of the original quadrats in permenanent livestock exclosures, \new{which we used} as control plots (i.e. ambient precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

@@

Ambient precipitation was variable over the five years of the study (Fig. 2A), meaning our experiment captured a range of environmental conditions.
ANPP varied from a minimum of 77.4 g m$^{-2}$ in 2014 to a maximum of 239.3 g m$^{-2}$ in 2016 when averaged across treatments (Fig. 2C).
ANPP was slightly higher in irrigation plots \new{(on average 15\% higher)} and slightly lower in drought plots \new{(on average 25\% lower) relative to control plots} (Fig. 2C), corresponding to observed and estimated soil volumetric water content (VWC) differences among treatments (Fig. 2B).
\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
Such differences in soil VWC indicate our treatment infrastructure was successful.
ANPP was highly variable across plots within years, \new{as indicated by the large and overlapping standard deviations} (Fig. 2C).

@@

\new{Sensitivity of ANPP to irrigation was constant over the course of our experiment (Fig. 3C). Sensivitity of ANPP to drought, however, grew over time (\emph{P} = 0.0007; Fig. 3C)}.

### 2018-01-05 manuscript/usses_water_ms.tex
big update with plug in VWC values from SOILWAT

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita},
\new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
perennial bunchgrasses, \emph{Pseudoroegneria spicata},
\emph{Poa secunda}, and \emph{Hesperostipa comata}
\new{(see Appendix 1 for rank abundance curves)}. During the period of
our experiment (2011 -- 2015), average mean annual precipitation was 265
mm year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
relocated 14 of the original quadrats in permenanent livestock
exclosures, \new{which we used} as control plots (i.e.~ambient
precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

@@

Ambient precipitation was variable over the five years of the study
(Fig. 2A), meaning our experiment captured a range of environmental
conditions. ANPP varied from a minimum of 74.5 g m\(^{-2}\) in 2014 to a
maximum of 237.1 g m\(^{-2}\) in 2016 when averaged across treatments
(Fig. 2C). ANPP was slightly higher in irrigation plots
\new{(on average 17\% higher)} and slightly lower in drought plots
\new{(on average 23\% lower) relative to control plots} (Fig. 2C),
corresponding to estimated soil volumetric water content (VWC)
differences among treatments (Fig. 2B).
\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
Such differences in soil VWC indicate our treatment infrastructure was
successful. ANPP was highly variable across plots within years,
\new{as indicated by the large and overlapping standard deviations}
(Fig. 2C).

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment ($\beta$ coefficients in equation 1). The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment. Posterior distributions of the offsets are in Figure 3A.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.16 & 0.14 & -1.22 & 1.69 \\ 
  Intercept & Drought & 0.13 & 0.05 & -1.69 & 2.39 \\ 
  Intercept & Irrigation & -0.24 & -0.18 & -2.11 & 1.53 \\ 
  Slope & Control & 0.94 & 0.86 & -0.51 & 2.69 \\ 
  Slope & Drought & 0.65 & 0.57 & -0.99 & 2.62 \\ 
  Slope & Irrigation & 0.72 & 0.65 & -0.54 & 2.26 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  30 &   2 & 1.92 & 0.01 \\ 
  2011 & betadisper &  30 &   2 & 0.64 & 0.53 \\ 
  2012 & adonis &  30 &   2 & 1.40 & 0.06 \\ 
  2012 & betadisper &  30 &   2 & 0.30 & 0.75 \\ 
  2013 & adonis &  30 &   2 & 1.70 & 0.01 \\ 
  2013 & betadisper &  30 &   2 & 0.59 & 0.56 \\ 
  2014 & adonis &  30 &   2 & 1.47 & 0.06 \\ 
  2014 & betadisper &  30 &   2 & 2.23 & 0.13 \\ 
  2015 & adonis &  30 &   2 & 1.22 & 0.14 \\ 
  2015 & betadisper &  30 &   2 & 1.78 & 0.19 \\ 
  2016 & adonis &  30 &   2 & 1.83 & 0.00 \\ 
  2016 & betadisper &  30 &   2 & 0.24 & 0.79 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}

>>>>

We conducted our precipitation manipulation experiment in a sagebrush
steppe community at the USDA, U.S. Sheep Experiment Station (USSES) near
Dubois, Idaho (44.2\(^{\circ}\) N, 112.1\(^{\circ}\) W), 1500 m above
sea level. The plant community is dominated by the shrub
\emph{Artemesia tripartita},
\new{the perennial forb \emph{Balsamorhiza sagittata}}, and three
perennial bunchgrasses, \emph{Pseudoroegneria spicata},
\emph{Poa secunda}, and \emph{Hesperostipa comata}
\new{(see Appendix 1 for rank abundance curves)}. During the period of
our experiment (2011 -- 2016), mean annual precipitation was 250 mm
year\(\phantom{}^{-1}\) and mean monthly temperature ranged from
-5.2\(^{\circ}\)C in January to 21.8\(^{\circ}\)C in July. Between 1926
and 1932, range scientists at the USSES established 26 permanent 1
m\(^2\) quadrats to track vegetation change over time. In 2007, we
relocated 14 of the original quadrats in permenanent livestock
exclosures, \new{which we used} as control plots (i.e.~ambient
precipitation) in the experiment described below.
\new{We used the relocated plots as our controls because collecting demographic data is difficult and it is already being collected in these plots for other ongoing studies.}

@@

Ambient precipitation was variable over the five years of the study
(Fig. 2A), meaning our experiment captured a range of environmental
conditions. ANPP varied from a minimum of 77.4 g m\(^{-2}\) in 2014 to a
maximum of 239.3 g m\(^{-2}\) in 2016 when averaged across treatments
(Fig. 2C). ANPP was slightly higher in irrigation plots
\new{(on average 15\% higher)} and slightly lower in drought plots
\new{(on average 25\% lower) relative to control plots} (Fig. 2C),
corresponding to observed and estimated soil volumetric water content
(VWC) differences among treatments (Fig. 2B).
\new{VWC in drought plots was 12\% less than in control plots on average, and VWC in irrigation plots was 19\% higher than in control plots on average.}
Such differences in soil VWC indicate our treatment infrastructure was
successful. ANPP was highly variable across plots within years,
\new{as indicated by the large and overlapping standard deviations}
(Fig. 2C).

@@

\begin{table}[ht]
\centering
\caption{Summary statistics from the posterior distributions of coefficients for each treatment ($\beta$ coefficients in equation 1). The `Intercept' and `Slope' summaries reported here for drought and irrigation are from the posterior distributions of the intercept and slope for the control treatment plus the offsets for each treatment. Posterior distributions of the offsets are in Figure 3A.} 
\begingroup\normalsize
\begin{tabular}{llrrrr}
  \hline
Coefficient & Treatment & Posterior Mean & Posterior Median & Lower 95\% BCI & Upper 95\% BCI \\ 
  \hline
Intercept & Control & 0.14 & 0.14 & -1.24 & 1.50 \\ 
  Intercept & Drought & 0.67 & 0.65 & -2.01 & 3.36 \\ 
  Intercept & Irrigation & -0.39 & -0.34 & -3.51 & 2.48 \\ 
  Slope & Control & 0.53 & 0.52 & -1.44 & 2.66 \\ 
  Slope & Drought & 1.02 & 1.00 & -1.04 & 3.25 \\ 
  Slope & Irrigation & 0.56 & 0.55 & -1.43 & 2.71 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}\begin{table}[ht]
\centering
\caption{Results from statistical tests for clustering and dispersion of community composition among precipitation treatments. `adonis' tests whether treatments form unique clusters in multidimensial space; `betadisper' tests whether treatments have similar dispersion. For both tests, \emph{P} values greater than 0.05 indicate there is no support that the treatments differ.} 
\begingroup\normalsize
\begin{tabular}{rlrrrr}
  \hline
Year & Test & n & d.f. & \emph{F} & \emph{P} \\ 
  \hline
2011 & adonis &  22 &   2 & 1.04 & 0.41 \\ 
  2011 & betadisper &  22 &   2 & 2.22 & 0.14 \\ 
  2012 & adonis &  22 &   2 & 1.09 & 0.32 \\ 
  2012 & betadisper &  22 &   2 & 0.21 & 0.81 \\ 
  2013 & adonis &  22 &   2 & 1.26 & 0.14 \\ 
  2013 & betadisper &  22 &   2 & 0.43 & 0.66 \\ 
  2014 & adonis &  22 &   2 & 0.96 & 0.52 \\ 
  2014 & betadisper &  22 &   2 & 0.36 & 0.70 \\ 
  2015 & adonis &  22 &   2 & 1.08 & 0.31 \\ 
  2015 & betadisper &  22 &   2 & 1.96 & 0.17 \\ 
  2016 & adonis &  22 &   2 & 1.12 & 0.27 \\ 
  2016 & betadisper &  22 &   2 & 0.27 & 0.76 \\ 
   \hline
\end{tabular}
\endgroup
\end{table}
